Variants on the Flight and Maintenance Planning problem

### 2018-08-28 latex/notes/model/model.tex

\usepackage[latin1]{inputenc}
% \usepackage{aeguill}
\usepackage[frenchb,english]{babel}
\usepackage{caption}

@@

  % TODO: previously assigned missions.

    \begin{tabular}{p{8mm}p{167mm}}
        $H_j$             & amount of resource time required by task $j$. \\
        $R_j$             & number of resources required by task $j$. \\
        $MT_j$            & minimum number of consecutive periods a resource can be assigned to task $j$. \\
        % $aut_i$           & maximal absolute usage time for resource $i$. \\
        % $aet_i$           & maximal absolute elapsed time for resource $i$. \\
        $M$               & maintenance duration in number of periods. \\
        $M^{max}$         & maximum number of simultaneous maintenance operations. \\
        % $E$               & remaining elapsed time after a maintenance. \\
        $E^{M}$         & maximum number of periods between two consecutive maintenances. \\
        $E^{m}$         & minimum number of periods between two consecutive maintenances. \\
        $H$               & remaining usage time after a maintenance. \\
        % $S$               & Storage duration in number of periods. \\
        % $W_1$             & weight of the first objective in the objective function. \\
        % $W_2$             & weight of the second objective in the objective function. \\
        $N_t$             & number of resources in already-planned maintenances in period $t$.\\
        % $D_t$             & number of resources to be assigned in total in period $t$. \\
        $AK_{kt}$         & maximum number of resources in cluster $k$ that can be simultaneously under maintenance at period $t$.\\
        $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning horizon. \\
        $Ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start of the planning horizon. \\
        % $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning horizon. \\
        % $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning horizon. \\
    \end{tabular}

@@

        \begin{tabular}{p{5mm}p{170mm}}
        $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which task $j$ is active. \\
        $\mathcal{J}_t $    &  tasks $j \in \mathcal{J}$ to be realized in period $t$. \\
        $\mathcal{I}_j$     &  resources $i \in \mathcal{I}$ that can be assigned to task $j$. \\
        $\mathcal{I}_k$     &  resources $i \in \mathcal{I}$ that are included in cluster $k$. \\
        $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$. \\
        % $\mathcal{T}^{st}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - S+1\}},  ..., {t}$\}. \\
        $\mathcal{T}^{m}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t}, ..., \min{\{|\mathcal{T}|, t + E^{m}-1\}}\}$. \\
        $\mathcal{T}^{M}_t$ &  time periods $t' \leq |\mathcal{T}| - E^{M}$ such that $t' \in \{ t + E^{m}-1 , ...,  t + E^{M}-1 \}$. \\
        $\mathcal{T}^{m_{ini}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ 1, ..., \max{\{0, Ret^{Init}_{i} - E^M + E^m \}}\}$. \\
        $\mathcal{T}^{M_{ini}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ \max{\{0, Ret^{Init}_{i} - E^M + E^m \}} , ...,  Ret^{Init}_{i} \}$. \\
        $\mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$. \\
        $\mathcal{A}^{init}_{jt}$  & resources $i \in \mathcal{I}$ that have task $j$ pre-assigned in period $t$. \\
    \end{tabular}

@@

    \begin{align}
        % if start task => assign start
        & a^s_{jti} \geq a_{jti} - a_{j(t-1)i} - 1 \!1_{i \in \mathcal{A}^{init}_{j(t-1)}}
                & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j \label{eq:start1} \\
      % # if we start a task, we need at least X periods of tasks assignments
        % & a^s_{jt'i} \leq a_{jti}
        % & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j, t' \in \mathcal{T}^{MT}_{jt} \label{eq:start2} \\
        % better constraint:
        & \sum_{t' \in \mathcal{T}^{MT}_{jt}} a^s_{jt'i} \leq a_{jti} 
        & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j \label{eq:start3}
    \end{align}

    TODO: explain constraint (\ref{eq:start3}).

    Constraints (\ref{eq:start1}) control the definition of starting a consecutive assignment: the variable is active if a resource $i$ is assigned a task $j$ in period $t$ and was not assigned this same task in the previous period ($t-1$). The previous period can include the period before the start of the planning horizon ($\mathcal{A}^{init}_{j(t-1)}$ set).
    Constraints (\ref{eq:start3}) control the minimum duration of a consecutive task assignment. If an assignment to task $j$ is started in period $t$, it needs to continue for the following $t' \in \mathcal{T}^{MT}_{jt}$ periods.

    \begin{align}
       & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}_k} m_{it'} \leq AK_{kt}
        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-cluster}
    \end{align}

    Constraints (\ref{eq:avalaibility-cluster}) guarantee a minimum availability for each cluster $k$ of resources. A cluster is defined by the biggest group of resources that is required, exclusively, by at least one task.

    \begin{align}
        % remaining used time
         & rut_{it} \leq rut_{i(t-1)} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
        & rut_{i0} = Rut^{Init}_i
               & i \in \mathcal{I} \label{eq:rut_initial}\\
        & rut_{it} \geq H m_{it}
                & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rut_lower}\\ 
        & rut_{it} \in [0,H]
                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu}              
        % & ret_{it} \leq ret_{it-1} - 1 + \sum_{t' \in \mathcal{T}^{st}_t} s_{it'} + E m_{it}
        %         & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:ret_upper}\\
        % & ret_{i0} = Ret^{Init}_i
        %         & i \in \mathcal{I} \label{eq:ret_initial}\\
        % & ret_{it} \geq E m_{it}
        %         & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_lower}\\                 
        % & ret_{it} \in [0,E]
        %         & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}\\
        % & \sum_{i \in \mathcal{I}} ret_{it} \geq Ret^{Init}_{sum}
        %       & t = |\mathcal{T}| \label{eq:min_ret}\\
        % & \sum_{i \in \mathcal{I}} rut_{i|\mathcal{T}|} \geq Rut^{Init}_{sum}\\
    \end{align}

>>>>

\usepackage[latin1]{inputenc}
% \usepackage{aeguill}
\usepackage[english]{babel}
\usepackage{caption}

@@

    \begin{tabular}{p{8mm}p{167mm}}
        $H_j$             & amount of resource time required by task $j$. \\
        $R_j$             & number of resources required by task $j$. \\
        $MT_j$            & minimum number of consecutive periods a resource can be assigned to task $j$. \\
        % $aut_i$           & maximal absolute usage time for resource $i$. \\
        % $aet_i$           & maximal absolute elapsed time for resource $i$. \\
        $M$               & maintenance duration in number of periods. \\
        $M^{max}$         & maximum number of simultaneous maintenance operations. \\
        % $E$               & remaining elapsed time after a maintenance. \\
        $E^{M}$         & maximum number of periods between two consecutive maintenances. \\
        $E^{m}$         & minimum number of periods between two consecutive maintenances. \\
        $H$               & remaining usage time after a maintenance. \\
        % $S$               & Storage duration in number of periods. \\
        % $W_1$             & weight of the first objective in the objective function. \\
        % $W_2$             & weight of the second objective in the objective function. \\
        $N_t$             & number of resources in already-planned maintenances in period $t$.\\
        $Nk_{kt}$         & number of resources in cluster $k$ in already-planned maintenances in period $t$.\\
        % $D_t$             & number of resources to be assigned in total in period $t$. \\
        $AK_{kt}$         & maximum number of resources in cluster $k$ that can be simultaneously under maintenance at period $t$.\\
        $HK_{kt}$         & minimum number of total remaining usage time for cluster $k$ at period $t$.\\
        $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning horizon. \\
        $Ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start of the planning horizon. \\
        % $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning horizon. \\
        % $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning horizon. \\
    \end{tabular}

@@

        \begin{tabular}{p{5mm}p{170mm}}
        $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which task $j$ is active. \\
        $\mathcal{J}_t $    &  tasks $j \in \mathcal{J}$ to be realized in period $t$. \\
        $\mathcal{I}_j$     &  resources $i \in \mathcal{I}$ that can be assigned to task $j$. \\
        $\mathcal{I}_k$     &  resources $i \in \mathcal{I}$ that are included in cluster $k$. One resource can belong to more than one cluster. \\
        $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$. \\
        % $\mathcal{T}^{st}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - S+1\}},  ..., {t}$\}. \\
        $\mathcal{T}^{m}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t}, ..., \min{\{|\mathcal{T}|, t + E^{m}-1\}}\}$. \\
        $\mathcal{T}^{M}_t$ &  time periods $t' \leq |\mathcal{T}| - E^{M}$ such that $t' \in \{ t + E^{m}-1 , ...,  t + E^{M}-1 \}$. \\
        $\mathcal{T}^{m_{ini}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ 1, ..., \max{\{0, Ret^{Init}_{i} - E^M + E^m \}}\}$. \\
        $\mathcal{T}^{M_{ini}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ \max{\{0, Ret^{Init}_{i} - E^M + E^m \}} , ...,  Ret^{Init}_{i} \}$. \\
        $\mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$. \\
        $\mathcal{A}^{init}_{j}$  & resources $i \in \mathcal{I}$ that have task $j$ pre-assigned in the previous period to the start of the horizon. \\
    \end{tabular}

@@

    \begin{align}
        % if start task => assign start
        & a^s_{jti} \geq a_{jti} - a_{j(t-1)i}
                & t =1, ..., \mathcal{T}, j \in \mathcal{J}_t, i \in \mathcal{I}_j \label{eq:start1} \\
        & a^s_{j0i} \geq a_{j0i} - 1 \!1_{i \in \mathcal{A}^{init}_{j}}
                & j \in \mathcal{J}_0, i \in \mathcal{I}_j \label{eq:start2} \\        
      % # if we start a task, we need at least X periods of tasks assignments
        % & a^s_{jt'i} \leq a_{jti}
        % & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j, t' \in \mathcal{T}^{MT}_{jt} \label{eq:start2} \\
        % better constraint:
        & \sum_{t' \in \mathcal{T}^{MT}_{jt}} a^s_{jt'i} \leq a_{jti} 
        & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j \label{eq:start3}
    \end{align}

    Constraints (\ref{eq:start1}) control the definition of starting a consecutive assignment: the variable is active if a resource $i$ is assigned a task $j$ in period $t$ and was not assigned this same task in the previous period ($t-1$). Constraints (\ref{eq:start2}) are the particular case for the first period on the horizon where, the previous assignments are given.

    Constraints (\ref{eq:start3}) control the minimum duration of a consecutive task assignment. If an assignment to task $j$ is started in period $t$, it needs to continue for the following $t' \in \mathcal{T}^{MT}_{jt}$ periods. They are a stronger version of the constraint $ a^s_{jt'i} \leq a_{jti}$.

    \begin{align}
       & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}_k} m_{it'} + Nk_{kt} \leq AK_{kt}
        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-cluster} \\
       & \sum_{i \in \mathcal{I}_k} rut_{it} \geq HK_{kt}
        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-hours-cluster}
    \end{align}

    Constraints (\ref{eq:avalaibility-cluster}) guarantee a minimum availability of resources for each cluster $k$. A cluster is defined by the biggest group of resources that is required, exclusively, by at least one task. Constraints (\ref{eq:avalaibility-hours-cluster}) guarantee that a minimum availability in terms of usage houres for each cluster $k$.

    \begin{align}
        % remaining used time
         & rut_{it} \leq rut_{i(t-1)} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j 
            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
        & rut_{i0} = Rut^{Init}_i
               & i \in \mathcal{I} \label{eq:rut_initial}\\
        & rut_{it} \geq H m_{it'}
                & t \in \mathcal{T}, t' \in \mathcal{T}^{s}_t, i \in \mathcal{I}\label{eq:rut_lower}\\ 
        & rut_{it} \in [0,H]
                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu}              
        % & ret_{it} \leq ret_{it-1} - 1 + \sum_{t' \in \mathcal{T}^{st}_t} s_{it'} + E m_{it}
        %         & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:ret_upper}\\
        % & ret_{i0} = Ret^{Init}_i
        %         & i \in \mathcal{I} \label{eq:ret_initial}\\
        % & ret_{it} \geq E m_{it}
        %         & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_lower}\\                 
        % & ret_{it} \in [0,E]
        %         & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}\\
        % & \sum_{i \in \mathcal{I}} ret_{it} \geq Ret^{Init}_{sum}
        %       & t = |\mathcal{T}| \label{eq:min_ret}\\
        % & \sum_{i \in \mathcal{I}} rut_{i|\mathcal{T}|} \geq Rut^{Init}_{sum}\\
    \end{align}
